\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-11 cc-condensed.tex
resolve comments

\noindent
Consistency says that the type $\bot$ is not inhabited.
Thus, it is safe to use the type theory for logical reasoning, as not
every proposition is trivially provable.
\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
the well-formedness rules, which, in particular, interprets $\bot$ in
\TT as an uninhabited type in the meta\-language.
The construction of this interpretation follows the standard model of
\citet{altkap2016}, extending it to handle linkages.
Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products,
as rule \ruleref{l/add} indicates.
%
We refer the reader to a supplemental file (in Agda syntax) for the complete proof.

@@

\noindent
This canonicity theorem says that every closed term of the ground
type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
When canonicity is proved in a constructive meta\-logic, its proof
amounts to a normalization function for closed terms of the ground type\EDJ{Suggesting use "boolean type". Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat). We also does specify what is ground type. }.
So canonicity serves to justify the computational nature of the type theory.

We prove \cref{thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
into a logical predicate on closed terms: the predicate includes all
closed, ``reducible''\EDJ{I will add a it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)} terms of type~$T$.
The proof is available as a supplemental file in Agda syntax.

>>>>

\noindent
Consistency says that the type $\bot$ is not inhabited.
Thus, it is safe to use the type theory for logical reasoning, as not
every proposition is trivially provable.

\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
the well-formedness rules, which, in particular, interprets $\bot$ in
\TT as an uninhabited type in the meta\-language.
The construction of this interpretation follows the standard model of
\citet{altkap2016}, extending it to handle linkages.
A linkage~$\LkgAdd{\lkg}{t}$ can be interpreted into a non-dependent pair where the
second component is a dependent function (modeling late binding),
%\EDJ{Suggesting changing to ``can be interpreted into non-dependent pair with dependent function for late-binding''}
as rule \ruleref{l/add} indicates.
%
We refer the reader to a supplemental file in Agda syntax for the complete proof.

@@

\noindent
This canonicity theorem says that every closed term of the ground
type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
When the canonicity theorem is proved in a constructive meta\-logic, its proof
amounts to a normalization function for closed terms of the ground type.
%\EDJ{Suggesting use "boolean type". Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat). We also does specify what is ground type. }\YZreply{I think it is clear from the context and from the article 'the' that 'the ground type' refers to B.}
So canonicity serves to justify the computational nature of the type theory.

We prove \cref{thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
into a logical predicate on closed terms: the predicate includes all
closed, ``reducible'' terms of type~$T$.
%\EDJ{I will add a quote around reducible here. it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}\EDJreply{Because in the operational semantic, using ``reducible'' is totally fine. But in declarative style, only ``canonicity'' this word is making sense. But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
The proof is available as a supplemental file in Agda syntax.

### 2022-11-11 metatheory-condensed.tex
resolve comments

The third premise
$\goodTerm{\Gm,\TyTkg{\lsig}}{\seal}{\sub{A}{\SubstWeak{1}}}$
in rules \ruleref{l/add} and \ruleref{lsig/add} is
responsible for hiding W-type signatures.
Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
that packages the types of all fields preceding the current field~$t$,
and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new tuple of type~$A$
that hides W-type signatures behind $\cU$, if necessary.
%
It is straightforward to find the~$\seal$ that fits the bill, though this process
is not automated in \TT.
In particular, when no hiding is needed, $\seal$ is $\var{0}$ (the identity)\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right? Or you meant to say it is identity ``sealing''?}\YZreply{Yes, I meant identity sealing. I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''} and $A$ is
$\TyTkg{\lsig}$.
In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are identities,
and $\seal_6$ is from
\begingroup
\fontsize{9}{9}\selectfont%
$\big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$
\endgroup
to
\begingroup
\fontsize{9}{9}\selectfont%
$\big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$%
\endgroup
.

>>>>

The third premise
$\goodTerm{\Gm,\TyTkg{\lsig}}{\seal}{\sub{A}{\SubstWeak{1}}}$
in rules \ruleref{l/add} and \ruleref{lsig/add} is
responsible for hiding W-type signatures.
Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
that packages the types of all fields preceding the current field~$t$,
and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new tuple of type~$A$
that hides W-type signatures behind $\cU$, if necessary.
%
It is straightforward to find the~$\seal$ that fits the bill, though this process
is not automated in \TT.
In particular, when no hiding is needed, $\seal$ is $\var{0}$
%\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right? Or you meant to say it is identity ``sealing''?}\YZreply{Yes, I meant identity sealing. I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''}
and $A$ is $\TyTkg{\lsig}$.
In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are $\var{0}$,
and $\seal_6$ is from
\begingroup
\fontsize{9}{9}\selectfont%
$\big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$
\endgroup
to
\begingroup
\fontsize{9}{9}\selectfont%
$\big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$%
\endgroup
.

### 2022-11-11 related-work.tex
resolve comments

On the other hand, the extra\-linguistic nature of the approaches tend
to lead to non-idiomatic code and offset their user-friendliness.
In particular, because data types have to be encoded (rather than expressed
through natively supported inductive types), the resulting code is
less readable and the resulting programming style inaccessible to non-experts.
In addition, extra programmer effort may be required, such as
having to manually prove additional well-formedness conditions on the algebras\EDJ{What is this `algebra'? I mean if it is a very frequent word to the audience  familiar with DTC then removing explanation is fine.}.

\citet{forsta2020} introduce \emph{Coq à la carte}.
It still follows DTC, but rather than embracing DTC's use of generic
fixed points, it considers specific instantiations instead.
The resulting mechanism appears more streamlined than prior \emph{à la
carte} approaches particularly for its extensive tool support for generating boilerplate code.
%
But even with the tool support, components of individually developed
feature extensions (e.g., "substlem") have to be composed separately by
invoking the tool.
%
%The approach does not seem to fully address the expression problem,
%as \citet{liwei2022} observe, in that further extending a composed
%mechanization causes (a small amount of) boilerplate code that glues definitions
%to be regenerated.

@@

Pumpkin Pi~\cite{ringer2021pumpkin} is a Coq plug-in that helps
repair proofs broken by changes in type definitions.
Its decompiler from proof terms to proof scripts prioritizes suggesting
useful tactics over soundness.
While Pumpkin Pi focuses on refactoring existing proofs in response to
changed definitions,
our solution can be viewed as an effort to preempt refactoring by
allowing the programmer to write code that has built-in hooks for
future extensions.

>>>>

On the other hand, the extra\-linguistic nature of the approaches tend
to lead to non-idiomatic code and offset their user-friendliness.
In particular, because data types have to be encoded (rather than expressed
through natively supported inductive types), the resulting code is
less readable and the resulting programming style inaccessible to non-experts.
In addition, extra programmer effort may be required, such as
having to manually prove additional well-formedness conditions. %on the algebras\EDJ{What is this `algebra'? I mean if it is a very frequent word to the audience  familiar with DTC then removing explanation is fine.}.

\citet{forsta2020} introduce \emph{Coq à la carte}.
It still follows DTC, but rather than embracing DTC's use of generic
fixed points, it considers specific instantiations instead.
The resulting mechanism appears more streamlined than prior \emph{à la
carte} approaches particularly for its extensive tool support for generating boilerplate code.
%
But even with the tool support, components (e.g., "substlem") of
individually developed feature extensions have to be composed separately
by invoking the tool.
%
%The approach does not seem to fully address the expression problem,
%as \citet{liwei2022} observe, in that further extending a composed
%mechanization causes (a small amount of) boilerplate code that glues definitions
%to be regenerated.

@@

Pumpkin Pi~\cite{ringer2021pumpkin} is a Coq plug-in that helps
repair proofs broken by changes in type definitions.
Its decompiler from proof terms to proof scripts prioritizes suggesting
useful tactics over soundness.
While Pumpkin Pi focuses on refactoring existing proofs in response to
changed definitions,
our solution can be viewed as an effort to preempt refactoring by
enabling the programmer to write code that has built-in hooks for
future extensions.
